Fast and scalable Gaussian process modeling with applications to astronomical time series \vspace{-3\baselineskip}

### 2017-01-04 paper/ms.tex

Another special case of the \genrp\ model of great physical interest is a
stochastically-driven simple harmonic oscillator.
The differential equation for this system is
\begin{equation}
    \left[\frac{\dd^2}{\dd t^2} + \frac{\omega_0}{Q}\,\frac{\dd}{\dd t}
    + \omega_0^2\right]\, y(t) = \epsilon(t)
\end{equation}
where $\omega_0$ is the frequency of the undamped oscillator, $Q$ is the
quality factor of the oscillator, and $\epsilon(t)$ is a stochastic driving
force.
In the limit of an infinite time series and white random forcing, the PSD of
this equation is given by \citep{Anderson:1990}
\begin{equation}\eqlabel{sho-psd}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0\,\omega_0^4}
    {(\omega^2-\omega_0^2)^2 + \omega_0^2\omega^2/Q^2}
\end{equation}
where $S_0$ is a normalization constant.
The power spectrum in \eq{sho-psd} matches \eq{genrp-psd} if
\begin{eqnarray}
a_j &=& S_0\,\omega_0\,Q \\
b_j &=& \frac{S_0\,\omega_0\,Q}{\sqrt{4\,Q^2-1}} \\
c_j &=& \frac{\omega_0}{2\,Q}\\
d_j &=& \frac{\omega_0}{2\,Q} \sqrt{4\,Q^2-1} \quad,
\end{eqnarray}
for $Q \ge \frac{1}{2}$.
For $0 < Q \le \frac{1}{2}$, \eq{sho-psd} can be captured by a pair of \genrp\
terms with parameters
\begin{eqnarray}
a_{j\pm} &=& \frac{1}{2}\,S_0\,\omega_0\,Q\,\left[ 1 \pm
        \frac{1}{\sqrt{1-4\,Q^2}}\right] \\
b_{j\pm} &=& 0 \\
    c_{j\pm} &=& \frac{\omega_0}{2\,Q}\,\left[1 \mp \sqrt{1-4\,Q^2}\right] \\
d_{j\pm} &=& 0 \quad.
\end{eqnarray}
It is interesting to note that, because of the damping, the characteristic
oscillation frequency in this model $d_j$, for any finite quality factor $Q$,
is not equal to the frequency of the undamped oscillator $\omega_0$.

>>>>

Another special case of the \genrp\ model of great physical interest is a
stochastically-driven simple harmonic oscillator.
The differential equation for this system is
\begin{equation}
    \left[\frac{\dd^2}{\dd t^2} + \frac{\omega_0}{Q}\,\frac{\dd}{\dd t}
    + \omega_0^2\right]\, y(t) = \epsilon(t)
\end{equation}
where $\omega_0$ is the frequency of the undamped oscillator, $Q$ is the
quality factor of the oscillator, and $\epsilon(t)$ is a stochastic driving
force.
In the limit of an infinite time series and white random forcing, the PSD of
this equation is given by \citep{Anderson:1990}
\begin{equation}\eqlabel{sho-psd}
S(\omega) = \sqrt{\frac{2}{\pi}}\,\frac{S_0\,\omega_0^4}
    {(\omega^2-\omega_0^2)^2 + \omega_0^2\omega^2/Q^2}
\end{equation}
where $S_0$ is a normalization constant.
The power spectrum in \eq{sho-psd} matches \eq{genrp-psd} if
\begin{eqnarray}
a_j &=& S_0\,\omega_0\,Q \\
b_j &=& \frac{S_0\,\omega_0\,Q}{\sqrt{4\,Q^2-1}} \\
c_j &=& \frac{\omega_0}{2\,Q}\\
d_j &=& \frac{\omega_0}{2\,Q} \sqrt{4\,Q^2-1} \quad,
\end{eqnarray}
for $Q \ge \frac{1}{2}$.
For $0 < Q \le \frac{1}{2}$, \eq{sho-psd} can be captured by a pair of \genrp\
terms with parameters
\begin{eqnarray}
a_{j\pm} &=& \frac{1}{2}\,S_0\,\omega_0\,Q\,\left[ 1 \pm
        \frac{1}{\sqrt{1-4\,Q^2}}\right] \\
b_{j\pm} &=& 0 \\
    c_{j\pm} &=& \frac{\omega_0}{2\,Q}\,\left[1 \mp \sqrt{1-4\,Q^2}\right] \\
d_{j\pm} &=& 0 \quad.
\end{eqnarray}

These identities yield a kernel of the form
\begin{equation}
k_{SHO}(\tau) = S_0 \omega_0 Q e^{-\frac{\omega_0 \tau}{2Q}} \begin{cases}
\cosh{(\eta \omega_0 \tau)} + (2Q\eta)^{-1} \sinh{(\eta \omega_0 \tau)}, & 0 < Q < 1/2\\
2 (1+\omega_0 \tau), & Q = 1/2\\
\cos{(\eta \omega_0 \tau)} + (2Q\eta)^{-1} \sin{(\eta \omega_0 \tau)},& 1/2 < Q < \infty\\
\end{cases}
\end{equation}
where $\eta = \vert 1-(4Q^2)^{-1}\vert$.
It is interesting to note that, because of the damping, the characteristic
oscillation frequency in this model $d_j$, for any finite quality factor $Q > 1/2$,
is not equal to the frequency of the undamped oscillator $\omega_0$.
